|

### 2022-05-28 memorandum/memo.Rmd
reduce word count article and memo

We thank Reviewer 1 for these two relevant suggestions. We now cite Gilligan and
Matsusaka's 2001 article three times in our paper. We mention their article
twice in our Introduction, first on page 2: ``\textit{On the one hand, enlarged
legislatures may reduce political particularism and increase overall government
spending, which can lead to more public goods  \citep{gerring2020population,
GilliganMatsusaka2001, Weingastetal1981}}'', and then on page 5: ``\textit{[The]
corollary of the "law of $1/n$" is that increasing legislature size promotes
excessive public spending, regardless of the partisan makeup of the legislature
\citep{GilliganMatsusaka2001}}''. We cite the article once more in our Theory
section (page 7): ``\textit{Yet, \citet{GilliganMatsusaka2001} find that in the
US legislature size is associated primarily with higher spending on
geographically concentrated benefits, such as education and highways}''.

We now cite Gerring and Veenendaal's book four times. We mention it twice on
page 2: ``\textit{Legislature size is known to have a significant effect on
collective action strategies, the bargaining power between state branches,
government spending, and representation \citep{crain1979cost,
freire2021legislature, gerring2020population, PrimoSnyder2008, rogers2002free,
Weingastetal1981}. However, how legislature size impacts social welfare remains
less understood. On the one hand, enlarged legislatures may reduce political
particularism and increase overall government spending, which can lead to more
public goods \citep{gerring2020population, GilliganMatsusaka2001,
Weingastetal1981}}''. We mention it again on page 4: ``\textit{Following the
seminal contribution by \citet{Weingastetal1981}, the "law of $1/n$" posits that
larger legislatures increase the legislative pressure for providing targeted
public goods to local supporters at the expense of the entire polity
\citep{freire2021legislature}. This question has become more relevant since
legislature size has increased consistently over time
\citep{gerring2020population}}''. Finally, we also cite the book on page 6:
``\textit{Larger legislatures improve representation as both the election
thresholds and the ratio of population-per-politician decrease, which facilitate
the electoral success of ethnic and political minorities
\citep{allenrepresentation, gerring2020population}}.''

@@

We thank Reviewer 1 for the opportunity to clarify why we have selected these
two sets of variables. We included an explanation for our decision on page 17:
"\textit{Our focus is on the impact of legislature size on education and health
care, which are the two most important public services the municipal government
should provide to citizens. According to the Brazilian Constitution,
municipalities must spend 25\% of their budget on education and 15\% on health
care. Consequently, the Ministry of Health and the Ministry of Education provide
readily available, fine-grained data on health care and education at the
municipal level, which we leverage in this study}."

@@

We agree with Reviewer 1 that we needed to provide more information about our
online survey. Therefore, we have detailed our survey procedures in Section A.14
of the Supplemental Materials (pages 23-24), and we also briefly discuss our
survey methodology on pages 17 and 22 (footnote 13) of the main article.
Furthermore, we have rewritten paragraph 2 of Section 6.2 to clarify the
importance of the survey results in our paper. It now reads as follows:
``\textit{We assume that city councilors want to extract rents and to get
reelected. The rents come from continuing in the repeated game, and the
partisanship discounts that benefit. Figure 2 shows that 65.9 percent of
councilors mentioned that mayors use job appointments to consolidate their
coalition support. Moreover, 59.0 percent of the surveyed councilors affirmed
that mayors use personal favors, such as shifting resources to places where the
councilors have electoral support, to solidify their governing coalition. This
provides further evidence that mayors do offer city councilors the benefits they
demand. In this regard, granting favors works as rent transfers for councilors.
Job appointments are also convenient for the mayors because patronage in Brazil
has positive results in terms of welfare \citep{toral2019benefits}.}''

@@

We thank Reviewer 1 for this constructive comment. The inclusion of control
variables in our model has two purposes. Firstly, the covariates improve the
consistency of our empirical results. In Section A.5 of the Appendix, we use
simulations to show that the inclusion of covariates enhances the consistency of
our RDD estimates. The simulations indicate that the controls improve
consistency even if we assume different distributions for the running variables
and for the treatment effects. Similaly, Section 5 of the manuscript
demonstrates that the covariates improve consistency in both the main estimates
and the placebo ones. Table 1 (page 20, reproduced below) shows that without
controls, we would overestimate the treatment effects (finding an increase of
1.63 legislators instead of an increase of exactly one legislator), and
underestimate the placebo treatment effects (finding a significant decrease of
0.31 legislators where there was supposed to be no effects in legislature size).
This substantial improvement is attributed to the addition of control variables,
which make the subpopulations in each threshold more comparable across all the
thresholds. Thus, we avoid spurious comparisons between cases on the left
(right) of one threshold and those on the right (left) of a different threshold.

\begin{table}[H] \centering 
  \caption{Research Design Validity Check (reproduction from the main paper)}
    \scalebox{0.9}{
  \begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
 & \multicolumn{4}{c}{\textbf{Panel A: Validity Check --- Treatment Effect on City Council Size}} \\ \cline{2-5}
   & \multicolumn{1}{c}{\small{(1)}} & 
 \multicolumn{1}{c}{\small{(2)}} & 
 \multicolumn{1}{c}{\small{(3)}} & 
 \multicolumn{1}{c}{\small{(4)}} \\

@@

To reflect these pieces of information, we have rewritten paragraph five in
Section 5 of the manuscript to (page 19):

``[...] \textit{we add
five controls to the estimates: Population in 2003; GDP per capita; the number
of seats in 2000; year; and a dummy for Northeast municipalities. The reason
for the inclusion of first variable is the multiple threshold assignments. We
add GDP per capita because wealthier cities tend to have better public
services. We include number of seats in 2000 because some municipalities could
have experienced a change provoked by the previous council size, thus
confounding our estimates. The Northeast dummy was added to control for the
fact that the Lula administration heavily invested in the region from 2003
until 2010, improving several development indicators \citep{de2015politica}. We
included an indicator for year to improve efficiency and control for seasonal
effects. Moreover, adding controls improves efficiency in RDD
\citep{calonico2019regression} and also enhances the consistency in the
multiple thresholds RDD.}''

@@

We agree with Reviewer 2 that testing the proposed mechanism is an exciting
project for a follow-up article. In the Discussion section (page 29), we write:
"\textit{Second, future research could evaluate whether additional legislators
help mayors gain access to federal or state transfers in countries where
resources are centrally provided, such as Japan. As suggested by
\citet{catalinac2020tournament}, in these situations the national government can
create a tournament between municipalities and provide resources to those which
offer more electoral support. Scholars might want to analyze whether this
tournament structure interacts with the mechanisms we suggest in this paper}."

@@

We collapse two concerns from Reviewer 3 since they both speak to core topics in
our paper. In Table 6 of the Supplemental Materials (page 21), we show that an
additional legislator usually does not lead to significant increases in public
expenditure. Therefore, our results speak more about the allocation of resources
than about government spending. In our section titled "Background on the
Brazilian Case" (page 14), we clarify why increasing the city council size often
results in a pro-mayor legislator. We write (page 15): "\textit{Brazil also
adopts a unique formula to allocate seats, which combines the Hare quota and the
D’Hondt method and strongly favors larger parties \citep{nicolau2015sistemas}.
This poses a major obstacle for small parties to reach the electoral quorum, so
they often join party coalitions to improve their chances of winning legislative
seats \citep[68]{Ames2001}. In Brazil, mayors are usually members of the largest
coalitions \citep{poulsen2018political}, thus there is an incentive for
candidates to side with the mayor before and after the election}." We mention
that information again on page 22, footnote 12: "\textit{The effect of city
council size on the number of mayor-aligned politicians stems from the rounding
rule adopted by the proportional representation system in Brazil. The rounding
rule favors the coalitions with more votes. Most of the time, the coalition of
the elected mayor is the one that receives more votes.}"

We did not find evidence that electing an additional legislator improves
descriptive representation in the city council. For instance, we see no
significant increase in the number of female or non-white legislators, as shown
in Table 4 (page 27). Finally, we show in the table below that the mayors'
parties are not driving our results. We have ommitted the table from the
Supplemental Materials to conserve space, but it is available on the article's
GitHub repository.

@@

We entirely agree with Reviewer 3 that we needed to discuss the scope conditions
of our findings in more detail. We acknowledge that the external validity of
our results might be conditional on the contextual characteristics of our
study. We have rewritten two paragraphs in the Discussion (page 29) to highligh
the implication of our results. We write: "\textit{Our findings have
implications that extend well beyond Brazil. A general form of our argument is
that it is easy for a politician to negotiate with members of her in-group
\citep{abbink2019, alt2018giving}. Specifically, several countries have
separate executive and legislative powers, such as the federal level of
presidential systems, state governments, and local city councils. As these
institutions are influenced by political allegiances, the dynamics we expose
here may be generalizable. Depending on legislative support, changing the
legislature size could increase the chance of electing politicians that are not
aligned with the executive, hence decreasing welfare. Despite the particular
effects on political support and welfare, the logic described in this paper
would remain relevant for empirically accessing the effect of enlarged
legislatures on welfare. Our findings may also help scholars understand why
other countries in Latin America and Asia can generate effective governance
despite having strong executives and large coalitions
\citep{pereira2012surprising}.}".

In the next paragraph (page 29), we indicate some of the limitations of our
findings and encourage scholars to conduct further research on the topic. We
write: "\textit{This paper opens several questions for future analysis. [...]
Third, scholars might want to investigate whether our findings remain valid
under different conditions. It is possible that the mechanisms we present here
are non-linear, with additional legislators having a positive or negative
impact on welfare depending on certain electoral characteristics. Finally, it
remains unclear how other legislative features affect citizen well-being. For
example, city councils may vary in size, monetary compensation for councilors,
committee structure, and internal power structure. Understanding how
legislature size interacts with other features would improve our knowledge
about how local and national legislatures impact public services.
These questions are crucial for institution design and welfare in developing
democracies.}"

>>>>

\begin{quote}
We thank Reviewer 1 for these two relevant suggestions. We now cite Gilligan
and Matsusaka's 2001 article three times in our paper. We mention their article
twice in our Introduction, first on page 2: ``\textit{Larger
legislatures may reduce political particularism and increase overall government
spending, which leads to more public goods  \citep{gerring2020population,
GilliganMatsusaka2001, Weingastetal1981}}'', and then on page 4:
``\textit{The corollary of the "law of $1/n$" is that increasing legislature
size promotes excessive public spending, regardless of the partisan makeup of
the legislature \citep{GilliganMatsusaka2001}}''. We cite the article once more
in our Theory section (page 6): ``\textit{Yet, \citet{GilliganMatsusaka2001}
find that in the US legislature size is associated primarily with higher
spending on geographically concentrated benefits, such as education and
highways}.''

We now cite Gerring and Veenendaal's book four times. We mention it twice on
page 2: ``\textit{Legislature size has a significant effect on collective action
strategies, the bargaining power between state branches, government spending,
and representation \citep{crain1979cost, freire2021legislature,
gerring2020population, PrimoSnyder2008, rogers2002free, Weingastetal1981}.
However, how legislature size impacts social welfare remains less understood.
Larger legislatures may reduce political particularism and increase overall
government spending, which leads to more public goods
\citep{gerring2020population, GilliganMatsusaka2001, Weingastetal1981}}.'' We
mention it again on page 4: ``\textit{Following the seminal contribution by
\citet{Weingastetal1981}, the "law of $1/n$" posits that larger legislatures
increase the legislative pressure for providing targeted public goods to local
supporters at the expense of the entire polity \citep{freire2021legislature}.
This question has become more relevant since legislature size has increased
consistently over time \citep{gerring2020population}}.'' Finally, we also cite
the book on page 5: ``\textit{On the one hand, larger legislatures improve
representation and facilitate the electoral success of ethnic and political
minorities\citep{allenrepresentation, gerring2020population}}.''

@@

We thank Reviewer 1 for the opportunity to clarify why we have selected these
two sets of variables. We included an explanation for our decision on pages
14-15: "\textit{Our focus is on the impact of legislature size on education and
health care, which are the two most important public services the municipal
government should provide to citizens. According to the Brazilian Constitution,
municipalities must spend 25\% of their budget on education and 15\% on health
care. Consequently, the Ministry of Health and the Ministry of Education
provide readily available, fine-grained data on health care and education at
the municipal level, which we leverage in this study}."

@@

We agree with Reviewer 1 that we needed to provide more information about our
online survey. Therefore, we have detailed our survey procedures in Section
A.14 of the Supplemental Materials (pages 23-24), and we also briefly discuss
our survey methodology on page 14 of the main article. Furthermore, we have
rewritten the second paragraph of Section 6.2 to clarify the importance of the
survey results in our paper. It now reads as follows: ``\textit{We assume that
city councilors want to extract rents and to get reelected. The rents come from
continuing in the repeated game, and the partisanship discounts that benefit.
Figure 2 shows that 65.9 percent of councilors mentioned that mayors use job
appointments to consolidate their coalition support. Moreover, 59.0 percent of
the surveyed councilors affirmed that mayors use personal favors, such as
shifting resources to places where the councilors have electoral support, to
solidify their governing coalition. This provides further evidence that mayors
do offer city councilors the benefits they demand. In this regard, granting
favors works as rent transfers for councilors. Job appointments are also
convenient for the mayors because patronage in Brazil has positive results in
terms of welfare \citep{toral2019benefits}.}''

@@

We thank Reviewer 1 for this constructive comment. The inclusion of control
variables in our model has two purposes. Firstly, the covariates improve the
consistency of our empirical results. In Section A.5 of the Appendix, we use
simulations to show that the inclusion of covariates enhances the consistency of
our RDD estimates. The simulations indicate that the controls improve
consistency even if we assume different distributions for the running variables
and for the treatment effects. Similaly, Section 5 of the manuscript
demonstrates that the covariates improve consistency in both the main estimates
and the placebo ones. Table \ref{validitycheck} (page 17, reproduced below) shows that without
controls, we would overestimate the treatment effects (finding an increase of
1.63 legislators instead of an increase of exactly one legislator), and
underestimate the placebo treatment effects (finding a significant decrease of
0.31 legislators where there was supposed to be no effects in legislature size).
This substantial improvement is attributed to the addition of control variables,
which make the subpopulations in each threshold more comparable across all the
thresholds. Thus, we avoid spurious comparisons between cases on the left
(right) of one threshold and those on the right (left) of a different threshold.

\begin{table}[H] \centering 
  \caption{Research Design Validity Check (reproduction from the main paper)}
    \label{validitycheck} 
    \scalebox{0.9}{
  \begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
 & \multicolumn{4}{c}{\textbf{Panel A: Validity Check --- Treatment Effect on City Council Size}} \\ \cline{2-5}
   & \multicolumn{1}{c}{\small{(1)}} & 
 \multicolumn{1}{c}{\small{(2)}} & 
 \multicolumn{1}{c}{\small{(3)}} & 
 \multicolumn{1}{c}{\small{(4)}} \\

@@

To reflect these pieces of information, we have rewritten paragraph five in
Section 5 of the manuscript (pages 16-17). It now reads:

``[...] \textit{we add five controls to the estimates: Population in 2003; GDP
per capita; the number of seats in 2000; year; and a dummy for Northeast
municipalities. The reason for the inclusion of first variable is the multiple
threshold assignments. We add GDP per capita because wealthier cities tend to
have better public services. We include the number of seats in 2000 because
some municipalities could have experienced a change provoked by the previous
council size, thus confounding our estimates. The Northeast dummy was added to
control for the fact that the Lula administration invested heavily in the
region from 2003 until 2010, improving several development indicators
\citep{de2015politica}. We included an indicator for year to improve efficiency
and control for seasonal effects. Moreover, adding controls improves efficiency
in RDD \citep{calonico2019regression} and also enhances the consistency in the
multiple thresholds RDD.}''

@@

We agree with Reviewer 2 that testing the proposed mechanism is an exciting
project for a follow-up article. In the Discussion section (page 26), we write:
"\textit{Second, future research could evaluate whether additional legislators
help mayors gain access to federal or state transfers in countries where
resources are centrally provided, such as Japan. As suggested by
\citet{catalinac2020tournament}, in these situations the national government can
create a tournament between municipalities and provide resources to those which
offer more electoral support. Scholars might want to analyze whether this
tournament structure interacts with the mechanisms we suggest in this paper}."

@@

We collapse two concerns from Reviewer 3 since they both speak to core topics in
our paper. In Table 6 of the Supplemental Materials (page 21), we show that an
additional legislator usually does not lead to significant increases in public
expenditure. Therefore, our results speak more about the allocation of resources
than about government spending. In our section titled "Background on the
Brazilian Case" (page 12), we clarify why increasing the city council size often
results in a pro-mayor legislator. We write (page 13): "\textit{Brazil also
adopts a unique formula to allocate seats, which combines the Hare quota and the
D’Hondt method and strongly favors larger parties \citep{nicolau2015sistemas}.
This poses a major obstacle for small parties to reach the electoral quorum, so
they often join electoral coalitions to improve their chances of winning legislative
seats \citep[68]{Ames2001}. In Brazil, mayors are usually members of the largest
coalitions \citep{poulsen2018political}, thus there is an incentive for
candidates to side with the mayor before and after the election}." We mention
that information again on page 19, footnote 3: "\textit{The effect of city
council size on the number of mayor-aligned politicians stems from the rounding
rule adopted by the proportional representation system in Brazil. The rounding
rule favors the coalitions with more votes. Most of the time, the coalition of
the elected mayor is the one that receives more votes.}"

We did not find evidence that electing an additional legislator improves
descriptive representation in the city council. For instance, we see no
significant increase in the number of female or non-white legislators, as shown
in Table 4 (page 24). Finally, we show in the table below that the mayors'
parties are not driving our results. We have ommitted the table from the
Supplemental Materials to conserve space, but it is available on the article's
GitHub repository.

@@

We entirely agree with Reviewer 3 that we needed to discuss the scope conditions
of our findings in more detail. We acknowledge that the external validity of
our results might be conditional on the contextual characteristics of our
study. We have rewritten two paragraphs in the Discussion (page 25) to highligh
the implication of our results. We write: ``\textit{Our findings have
implications that extend well beyond Brazil. A general form of our argument is
that it is easy for a politician to negotiate with members of her in-group
\citep{abbink2019, alt2018giving}. Specifically, several countries have
separate executive and legislative powers, such as the federal level of
presidential systems, state governments, and local city councils. As these
institutions are influenced by political allegiances, the dynamics we expose
here may be generalizable. Depending on legislative support, changing the
legislature size could increase the chance of electing politicians that are not
aligned with the executive, hence decreasing welfare. Despite the particular
effects on political support and welfare, the logic described in this paper
would remain relevant for empirically accessing the effect of enlarged
legislatures on welfare. Our findings may also help scholars understand why
other countries in Latin America and Asia can generate effective governance
despite having strong executives and large coalitions
\citep{pereira2012surprising}.}''

In the next paragraph (page 26), we indicate some of the limitations of our
findings and encourage scholars to conduct further research on the topic. We
write: ``\textit{This paper opens several questions for future analysis. [...]
Third, scholars might want to investigate whether our findings remain valid
under different conditions. It is possible that the mechanisms we present here
are non-linear, with additional legislators having a positive or negative
impact on welfare depending on certain electoral characteristics. Finally, it
remains unclear how other legislative features affect citizen well-being. For
example, city councils may vary in size, monetary compensation for councilors,
committee structure, and internal power structure. Understanding how
legislature size interacts with other features would improve our knowledge
about how local and national legislatures impact public services.
These questions are crucial for institution design and welfare in developing
democracies.}''
